---
title: 'Reproducible Research: Peer Assessment 1'
author: "Leonardo Becerra L."
date: "Sunday, January 18, 2015"
output:
  html_document:
    keep_md: yes
---

## Loading and preprocessing the data

###Loading required packages

```{r}
require(plyr)
require(dplyr)
require(lubridate)
require(ggplot2)
require(ggthemes)
```
The first step is import the base data *activity,csv* where the original file incorporate headers and missing values are coded as NA.  
It's additionaly important to avoid the import to convert to Factors any data, and convert dates to date format.

```{r loading date, echo = TRUE}
Data <- read.csv("activity.csv", header = TRUE, na.strings = NA, stringsAsFactors = FALSE)
Data$date <- strptime(Data$date, format = "%Y-%m-%d")
# Verifing properties of fields
str(Data)
```


## What is mean total number of steps taken per day?

### Extracting first day and month of each Date, to be able to filter and graph variable for month and day.
```{r, echo = TRUE}
Data$month <- months(Data$date)
Data$day <- yday(Data$date)
```

Selecting the required data froma Data  
Grouping by day to summarize statistics by each day
```{r, echo = TRUE}
per_day <- select(Data, steps, day, month) %>%
        group_by(day, month) %>% 
        summarize(sum = sum(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
```

Reporting a histogram of the mean total number of steps taken each day

```{r}
p <- ggplot(per_day, aes(sum, fill = month))
p + geom_histogram() + ggtitle("Mean total steps per day") + xlab("Mean steps per day")
```

Genratign and calculating the summary for mean and median numbers per day
```{r}
mean  <- mean(per_day$sum, na.rm = TRUE)
median  <- median(per_day$sum, na.rm = TRUE)
summary(per_day)
```

The mean value of steps per day is `r mean`  

The median value of steps per day is `r median`  

## Are there differences in activity patterns between weekdays and weekends?
```{r}
per_day_interval <- select(Data, steps, day, interval) %>%
        group_by(interval) %>% 
        summarize(mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
```

Reporting a time series plot of the mean total number of steps taken each day

```{r}
q <- ggplot(per_day_interval, aes(x = interval, y = mean))
q + geom_line() + ggtitle("Daily activity pattern") + xlab("interval") + ylab("mean steps per day")
```

Genratign and calculating the summary for mean and median numbers per day
```{r}
mean_perinterval  <- mean(per_day_interval$mean, na.rm = TRUE)
median_perinterval  <- median(per_day_interval$median, na.rm = TRUE)
summary(per_day)
```

The mean value of steps per day is `r mean_perinterval`  

The median value of steps per day is `r median_perinterval` 
## What is mean total number of steps taken per day?
```{r, echo = TRUE}
```

## What is the average daily activity pattern?

Defining if 
```{r}
Data$wday = as.POSIXlt(Data$date)$wday
Data$weekday <- wday(Data$date, label = TRUE)
tail(Data, 10)
Data$is_wkday <- ifelse(Data$wday %in% c("6", "0"),"weekend", "weekday")
Data$is_wkday <- as.factor(Data$is_wkday)
str(Data)
```

```{r}
per_day_interval_wday <- select(Data, steps, day, interval, is_wkday) %>%
        group_by(interval, is_wkday) %>% 
        summarize(mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
```

Reporting a time series plot of the mean total number of steps taken each day

```{r}
q <- ggplot(per_day_interval_wday, aes(x = interval, y = mean))
q + geom_line() + facet_wrap(~is_wkday) + ggtitle("Daily activity pattern") + xlab("interval") + ylab("mean steps per day")
```

Genratign and calculating the summary for mean and median numbers per day
```{r}
mean_perinterval  <- mean(per_day_interval$mean, na.rm = TRUE)
median_perinterval  <- median(per_day_interval$median, na.rm = TRUE)
summary(per_day)
```

The mean value of steps per day is `r mean_perinterval`  

The median value of steps per day is `r median_perinterval` 